---
title: "Creating beautiful tables with `{flextable}`"
format: 
  revealjs:
    theme: custom_rrus.scss
    menu: false
    progress: true
    slide-number: true
    show-slide-number: print
    center: true # This is my personal style (will remove if not okay)
execute: 
  message: false
  warning: false
title-slide-attributes:
  data-background-image: "https://raw.githubusercontent.com/rfortherestofus/rfortherestofus/main/inst/assets/rru-hex-bg-gradient-dark.svg"
  data-background-size: cover
---


## About me
```{r}
# Overwrite flextable output to export table as image and
# include it as image
knit_print.flextable <- function(x, options) {
  # Requires tbl.cap (not tbl-cap as is usual in Quarto)
  full_name <- paste0('img/', options$tbl.cap, '.png')
  x |> 
    flextable::save_as_image(path = full_name, webshot = 'webshot2')
  
  knitr::include_graphics(full_name)
}

registerS3method(
  "knit_print", 'flextable', knit_print.flextable, 
  envir = asNamespace("flextable") 
  # important to overwrite {flextable}s knit_print
)

```


::: notes
I'll introduce myself here.
2 min tops.
:::

## Where I've come from

::: panel-tabset
### Data

```{r}
#| echo: true
library(tidyverse)
penguins <- palmerpenguins::penguins |> 
  filter(!is.na(sex))
penguins
```

### My bad table

Bad table that I would have created in the past.
:::

## Avoid vertical lines with `flextable()`

```{r}
penguin_counts <- penguins |> 
  mutate(year = as.character(year)) |> 
  group_by(species, island, sex, year) |> 
  summarise(n = n(), .groups = 'drop')


penguin_counts_wider <- penguin_counts |> 
  pivot_wider(
    names_from = c(species, sex),
    values_from = n
  ) |> 
  arrange(island, year) 
knitr::opts_chunk$set("ft.shadow" = FALSE)
```

```{r}
#| tbl.cap: first_flextable

library(flextable)
library(officer) # Load as well for later theming.

flextable(penguin_counts_wider) |> 
  autofit()
```

## Use better column names and a title

```{r}
#| tbl.cap: better_name_flextable
new_names <- c('Island', 'Year', rep(c('Female', 'Male'), 3))
names(new_names) <- colnames(penguin_counts_wider)

flextable(penguin_counts_wider) |> 
  set_header_labels(
    values = new_names
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(2, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = c('Penguins in the Palmer Archipelago', 'Data is courtesy of the {palmerpenguins} R package')
  ) |> 
  autofit()
```

## Align columns

```{r}
#| tbl.cap: columns_aligned
flextable(penguin_counts_wider) |> 
  set_header_labels(
    island = 'Island',
    year = 'Year',
    Adelie_female = 'Female',
    Adelie_male = 'Male',
    Chinstrap_female = 'Female',
    Chinstrap_male = 'Male',
    Gentoo_female = 'Female',
    Gentoo_male = 'Male'
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(2, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = c('Penguins in the Palmer Archipelago', 'Data is courtesy of the {palmerpenguins} R package')
  ) |> 
  align(i = 3, align = 'center', part = 'header') |> 
  # it's already left-aligned but for demo purposes we do it again
  align(j = 'year', align = 'right', part = 'body') |> 
  autofit()
```

## Use groups instead of repetitive columns

```{r}
#| tbl.cap: grouped_columns
penguin_counts_wider |> 
  mutate(island = paste('Island: ', island)) |> 
  as_grouped_data(
    groups = 'island'
  ) |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  set_header_labels(
    island = 'Island',
    year = '',
    Adelie_female = 'Female',
    Adelie_male = 'Male',
    Chinstrap_female = 'Female',
    Chinstrap_male = 'Male',
    Gentoo_female = 'Female',
    Gentoo_male = 'Male'
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(1, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = c('Penguins in the Palmer Archipelago', 'Data is courtesy of the {palmerpenguins} R package')
  ) |> 
  align(i = 3, align = 'center', part = 'header') |> 
  autofit()
```

::: notes
This lets us also remove the `year` label
:::

## Format missing numbers

```{r}
#| tbl.cap: format_missing_numbers
penguin_counts_wider |> 
  mutate(island = paste('Island: ', island)) |> 
  as_grouped_data(
    groups = 'island'
  ) |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  set_header_labels(
    island = 'Island',
    year = '',
    Adelie_female = 'Female',
    Adelie_male = 'Male',
    Chinstrap_female = 'Female',
    Chinstrap_male = 'Male',
    Gentoo_female = 'Female',
    Gentoo_male = 'Male'
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(1, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = 'Penguins in the Palmer Archipelago\nData is courtesy of the {palmerpenguins} R package'
  ) |> 
  align(i = 2, align = 'center', part = 'header') |> 
  colformat_num(i = ~ (is.na(island)), na_str = '-') |> 
  autofit()
```

## Add summaries

```{r}
#| tbl.cap: add_summaries

maximum_summary <- penguin_counts_wider |> 
  group_by(island) |> 
  summarise(across(
    .cols = -year,
    .fns = max
  )) |> 
  mutate(year = 'Maximum')

total_summary <- penguin_counts_wider |> 
  group_by(island) |> 
  summarise(across(
    .cols = -year,
    .fns = sum
  )) |> 
  mutate(year = 'Total')
  

bind_rows(penguin_counts_wider, maximum_summary, total_summary) |> 
  mutate(island = paste('Island: ', island)) |> 
  arrange(island, year) |> 
  as_grouped_data(groups = 'island') |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  set_header_labels(
    island = 'Island',
    year = '',
    Adelie_female = 'Female',
    Adelie_male = 'Male',
    Chinstrap_female = 'Female',
    Chinstrap_male = 'Male',
    Gentoo_female = 'Female',
    Gentoo_male = 'Male'
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(1, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = c('Penguins in the Palmer Archipelago', 'Data is courtesy of the {palmerpenguins} R package')
  ) |> 
  align(i = 3, align = 'center', part = 'header') |> 
  colformat_num(i = ~ (is.na(island)), na_str = '-') |> 
  align(
    i = ~ (year %in% 2007:2009),
    j = 'year',
    align = 'right'
  ) |> 
  autofit()
```

## Style your table

```{r}
#| tbl.cap: adjust_cell_properties
final_table <- bind_rows(penguin_counts_wider, maximum_summary, total_summary) |> 
  arrange(island, year) |> 
  mutate(island = paste('Island: ', island)) |> 
  as_grouped_data(groups = 'island') |> 
  as_flextable(hide_grouplabel = TRUE) |> 
  set_header_labels(
    island = 'Island',
    year = '',
    Adelie_female = 'Female',
    Adelie_male = 'Male',
    Chinstrap_female = 'Female',
    Chinstrap_male = 'Male',
    Gentoo_female = 'Female',
    Gentoo_male = 'Male'
  ) |> 
  add_header_row(
    values = c('', 'Adelie', 'Chinstrap', 'Gentoo'),
    colwidths = c(1, 2, 2, 2)
  ) |> 
  add_header_lines(
    values = c('Penguins in the Palmer Archipelago', 'Data is courtesy of the {palmerpenguins} R package')
  ) |> 
  align(i = 3, align = 'center', part = 'header') |> 
  colformat_num(i = ~ (is.na(island)), na_str = '-') |> 
  align(
    i = ~ (year %in% 2007:2009),
    j = 'year',
    align = 'right'
  ) |> 
  hline(i = ~ (!is.na(island) | year %in% c('2009', 'Total')), border = fp_border(width = 1, color = 'grey40')) |> 
  bg(i = 3:4, bg = 'dodgerblue4', part = 'header') |> 
  bg(i = ~ (year %in% seq(2007, 2009, 2)), bg = 'grey90') |> 
  bold(part = 'header', i = 3) |> 
  color(i = 3:4, part = 'header', color = 'white') |> 
  border(j = 'year', border.right = fp_border(color = 'grey40', width = 1)) |> 
  style( #### subtitle
    part = 'header',
    i = 2,
    pr_t = fp_text(font.size = 11),
    ### This part is added 
    pr_c = fp_cell(border = fp_border(width = 0, color = 'grey40'))
  ) |>
  style( ### title
    part = 'header',
    i = 1,
    pr_t = fp_text(
      color = 'dodgerblue4',
      font.size = 18,
      font.family = 'Merriweather',
      bold = TRUE
    ),
    ### This part is added 
    pr_c = fp_cell(
      border.top = fp_border(width = 1.5, color = 'grey40'),
      border.bottom = fp_border(width = 0, color = 'grey40')
    )
  ) |> 
  style( ### group rows
    i = ~ (!is.na(island)),
    ### This part is added 
    pr_c = fp_cell(
      border.top = fp_border(width = 1, color = 'grey40'),
      border.bottom = fp_border(width = 1, color = 'grey40'),
      border.right = fp_border(width = 0, color = 'grey40')
    ),
    pr_t = fp_text(
      color = 'dodgerblue4',
      bold = TRUE,
      font.size = 12,
    )
  ) |> 
  line_spacing(part= 'all', space = 0.9) |> 
  ### This part is added 
  hline(part = 'header', i = 2, border = fp_border(width = 0)) |> 
  hline(part = 'header', i = 3, border = fp_border(width = 0)) |> 
  autofit()
final_table
```

::: notes
-   Horizontal lines for more structure
-   Background colors
-   Text appearance (font, color, size)
-   Adjust cell properties (some borders)
:::

# Sample Inverse {.inverse}

## Sample My Turn {.my-turn}

Exercitation consectetur sit excepteur cillum ad irure labore pariatur incididunt.
Esse enim mollit velit nulla eiusmod consectetur occaecat laborum excepteur eu laborum enim pariatur do.
Eiusmod exercitation minim incididunt nisi Lorem.
Consectetur irure laboris est id aliquip Lorem.
Minim et minim tempor ut aute minim.
Velit non officia nisi proident enim sint.
Nisi irure eu reprehenderit duis laboris amet velit velit consectetur magna quis.
Veniam quis incididunt do non ipsum.
Ut deserunt dolore sint aliquip ex deserunt eiusmod cillum ut.
Qui sit officia in voluptate officia qui enim eiusmod id dolor incididunt est.
